---
title: "Final Capstone Project"
author: "Moira Lennox"
date: "November 2r, 2015"
output: pdf_document
---

```{r setup,echo=FALSE} 
setwd("/Users/moiralennox/GitHub/Capstone")
#setwd("~/Documents/IT Documents/Data Science Class/DSCapStone")

# get packages that are needed
library(knitr)
library(data.table)
library(car) 
library(ggplot2)

```

## Title & Introduction
The data set for this capstone project was supplied by Yelp and is a subset of data for the years 2004 – 2015 and contains a number of counties. The yelp data supplied was made up of 5 files namely: reviews, business, user, tips, checkin. My work will focus on two files namely the review and business files.

My objective for this report is to analyze the “yelp” data set by exploring and researching the relationship between sets of variables and the business attribute that identifies as “dog allowed” to answer some questions. 

Dogs can be a major part of the family. We love taking our dog everywhere with us, however she is not always welcome. So here is my question: Can the yelp dataset tell me if there was a trend in welcoming pets into various businesses? I plan to narrow my focus down to two major categories hotels and restaurants since these are the areas I care about the most. I am going to look at the review data from 2008-2014. I will compare business that allow dogs to business that do not and sum the review count for year over year comparison and then try to correlate those counts to dog references I find from data mining the review text for those references. I will look at the country trends to see if locations makes a difference.

There is an attribute on the business dataset that determines if Dogs are allowed in that organization.


## Data processing and Transformation
I read in the json files for Business and Review, flatten them created a few new columns: main category, country
and review year. I aggregated the review data up to the business level. The final step was merge the dataset into a single dataset. 


```{r setup2 ,echo=FALSE}

load(file = "dt_merge.rda")
#subset to year 2008-2015
dt_merge_pet <- subset(dt_merge, revyear!=2015 & revyear!=2004 & revyear!=2005 & revyear!=2006 & revyear!=2007)
dt_merge_pet$revyear = factor(dt_merge_pet$revyear)
dt_merge_pet$country = factor(dt_merge_pet$country)


dt_merge_pet$petlog <- log(dt_merge_pet$petcount+1)
dt_merge_pet$revlog <- log(dt_merge_pet$revcount+1)
dt_merge_pet$votefunlog <- log(dt_merge_pet$votefuncount+1)
dt_merge_pet$voteusefullog <- log(dt_merge_pet$voteusefulcount+1)
dt_merge_pet$votecoollog <- log(dt_merge_pet$votecoolcount+1)

# Create just for box plot
dt_merge_pet2 <- subset(dt_merge_pet, dogsallowed== TRUE &revyear!=2015 & revyear!=2004 & revyear!=2005 & revyear!=2006 & revyear!=2007)


```

```{r setup3 ,echo=TRUE}
str(dt_merge_pet)

```

## Exploratory Data Analysis
A quick boxplot to review the pet count per year. The data is heavyly
skewed so we I did the log of the petcount.

```{r da1,echo=FALSE, warnings = FALSE}
ggplot(data=dt_merge_pet2, aes(x=dogsallowed, y=petlog, fill=dogsallowed)) + geom_boxplot() +
    stat_summary(fun.y=mean, colour="darkred", geom="point", 
                 shape=18, size=3,show_guide = FALSE) + 
    facet_grid(.~revyear) +
    ggtitle("Log of Pet Count by Year") +
    labs(x="Year",y="Log of Pet Count") 

```

```{r da2,echo=FALSE, warnings = FALSE}
means <- aggregate(petlog~revyear, dt_merge_pet2, mean)
means$petlog <- format(round(means$petlog, 2), nsmall = 2)
print(t(means))

```

```{r da3,echo=FALSE}

# create some new function to create scatter plot
panel.cor <- function(x, y, digits=2, cex.cor)
{
    usr <- par("usr"); on.exit(par(usr))
    par(usr = c(0, 1, 0, 1))
    r <- abs(cor(x, y))
    txt <- format(c(r, 0.123456789), digits=digits)[1]
    test <- cor.test(x,y)
    Signif <- ifelse(round(test$p.value,3)<0.001,"p<0.001",paste("p=",round(test$p.value,3)))  
    text(0.5, 0.25, paste("r=",txt))
    text(.5, .75, Signif)
}

panel.smooth<-function (x, y, col = "blue", bg = NA, pch = 18, 
                        cex = 0.8, col.smooth = "red", span = 2/3, iter = 3, ...) 
{
    points(x, y, pch = pch, col = col, bg = bg, cex = cex)
    ok <- is.finite(x) & is.finite(y)
    if (any(ok)) 
        lines(stats::lowess(x[ok], y[ok], f = span, iter = iter), 
              col = col.smooth, ...)
}

panel.hist <- function(x, ...)
{
    usr <- par("usr"); on.exit(par(usr))
    par(usr = c(usr[1:2], 0, 1.5) )
    h <- hist(x, plot = FALSE)
    breaks <- h$breaks; nB <- length(breaks)
    y <- h$counts; y <- y/max(y)
    rect(breaks[-nB], 0, breaks[-1], y, col="cyan", ...)
}

```


```{r da4,echo=FALSE}

pairs(~dogsallowed+petcount+revcount+revyear+maincat+country+votefuncount+voteusefulcount+votecoolcount+starsavg, 
      data=dt_merge_pet,
      lower.panel=panel.smooth, 
      upper.panel=panel.cor,
      diag.panel=panel.hist,
      main="Fig 2. Scatterplot Matrix")

```


## Methods and Data
With the use of linear regression models I quantify the relationships seen in the Exploratory Data Analysis section to find the best model fit and compare it with the base model.

```{r rm1,echo=FALSE}

fit<- lm(petcount~dogsallowed+revcount+revyear+maincat+country+votefuncount+
         voteusefulcount+votecoolcount+starsavg-1,data=dt_merge_pet)
summary(fit)


fit<- lm(petcount~revcount+votefuncount+voteusefulcount+votecoolcount-1,data=dt_merge_pet)
summary(fit)


# Dogs true
fit2<- lm(petlog~revlog+revyear+maincat+country+votefunlog+voteusefullog+votecoollog+starsavg -1,
          data=dt_merge_pet2)
summary(fit2)

fit2<- lm(petlog~revlog+maincat+votefunlog+country+votecoollog-1,data=dt_merge_pet2)
summary(fit2)

```


## Results


## Discussion & Conclusion


## Appendix
Below is the code the was used to create my data set which I will use for all my anaylsis.

This section will create 3 new functions below.

```{r createnewfun, eval=FALSE }
# create some new function to create scatter plot
panel.cor <- function(x, y, digits=2, cex.cor)
{
    usr <- par("usr"); on.exit(par(usr))
    par(usr = c(0, 1, 0, 1))
    r <- abs(cor(x, y))
    txt <- format(c(r, 0.123456789), digits=digits)[1]
    test <- cor.test(x,y)
    Signif <- ifelse(round(test$p.value,3)<0.001,"p<0.001",paste("p=",round(test$p.value,3)))  
    text(0.5, 0.25, paste("r=",txt))
    text(.5, .75, Signif)
}

panel.smooth<-function (x, y, col = "blue", bg = NA, pch = 18, 
                        cex = 0.8, col.smooth = "red", span = 2/3, iter = 3, ...) 
{
    points(x, y, pch = pch, col = col, bg = bg, cex = cex)
    ok <- is.finite(x) & is.finite(y)
    if (any(ok)) 
        lines(stats::lowess(x[ok], y[ok], f = span, iter = iter), 
              col = col.smooth, ...)
}

panel.hist <- function(x, ...)
{
    usr <- par("usr"); on.exit(par(usr))
    par(usr = c(usr[1:2], 0, 1.5) )
    h <- hist(x, plot = FALSE)
    breaks <- h$breaks; nB <- length(breaks)
    y <- h$counts; y <- y/max(y)
    rect(breaks[-nB], 0, breaks[-1], y, col="cyan", ...)
}


```

